---
title: "A Universal Saver's Match"
subtitle: "Design, reach, generosity, and cost of a redesigned federal retirement-savings match"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
params:
  root: NULL
output:
  html_document:
    self_contained: true
    theme: null
    highlight: null
    mathjax: null
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      results = "asis", fig.align = "center")

suppressPackageStartupMessages({
  library(arrow)
  library(dplyr)
  library(readr)
  library(knitr)
  library(kableExtra)
})

## ---- Resolve the project root (param -> env var -> upward search) -----------
root <- params$root
if (is.null(root) || !nzchar(root) || !dir.exists(root)) {
  root <- Sys.getenv("EIG_PROJECT_ROOT", unset = "")
}
if (!nzchar(root) || !dir.exists(root)) {
  cand <- normalizePath(getwd())
  while (!dir.exists(file.path(cand, "Infrastructure")) && cand != dirname(cand)) {
    cand <- dirname(cand)
  }
  root <- cand
}

dp <- file.path(root, "data", "processed", "universal_sm_hybrid")
rep_dir <- file.path(root, "output", "reports", "universal_sm_hybrid")
fig_dir <- file.path(root, "output", "figures", "universal_sm_hybrid")

## ---- Load the live pipeline artifacts ---------------------------------------
scen   <- as.data.frame(read_parquet(file.path(dp, "scenario_results.parquet")))
pivots <- readRDS(file.path(dp, "pivot_table.rds"))$pivot_tbl
sim    <- as.data.frame(read_parquet(file.path(dp, "simulation_results.parquet")))
funnel <- read_csv(file.path(rep_dir, "universe_funnel.csv"), show_col_types = FALSE)

get_scen <- function(nm) scen[scen$scenario_name_chr == nm, , drop = FALSE][1, ]
# Headline level of interest: automatic enrollment at 80 percent participation.
# The auto-enrollment literature (Madrian-Shea; state auto-IRA programs) and the
# review panel both judge the SIPP-observed conditional rate (~59 percent) to
# understate participation under a sticky payroll default, so the 80 percent
# auto-enrollment case is the central estimate; the observed-conditional case is
# carried as a behavioral lower bound and full participation as the upper bound.
headline <- get_scen("sens_auto_enroll_80pct")
obs_cond <- get_scen("headline_sipp_observed_conditional")
s_full   <- get_scen("sens_full_participation_100pct")

pv <- function(grp, col) pivots[[col]][pivots$filing_group_chr == grp]

## ---- Current-law Saver's Match contrast (same population, full take-up) ------
cl_avail <- FALSE
cl_path <- file.path(root, "output", "tables", "main", "sm_jct_replication_scenarios.xlsx")
if (requireNamespace("openxlsx", quietly = TRUE) && file.exists(cl_path)) {
  cl <- openxlsx::read.xlsx(cl_path, sheet = "All Scenario Results")
  cl_row <- cl[cl$Scenario == "universal_m100", , drop = FALSE]
  if (nrow(cl_row) == 1L) {
    cl_avail   <- TRUE
    cl_elig_M  <- cl_row[["Eligible.(M)"]]
    cl_cost_M  <- cl_row[["Annual.Cost.($M)"]]
    cl_avg     <- cl_row[["Avg.Match/Person.($)"]]
  }
}

## ---- Universe / sample funnel -----------------------------------------------
fn <- function(step) funnel$weighted_n_num[funnel$step_label_chr == step] / 1e6
adults_18p_M <- fn("02_age_18_plus")
earners_M    <- fn("05_has_annual_earned_income")
universe_M   <- fn("06_valid_filing_group")

## ---- Distributional incidence (full-participation match dollars) ------------
inc <- sim[!is.na(sim$magi_num), ]
inc <- inc[order(inc$magi_num), ]
inc$cw  <- cumsum(inc$WPFINWGT)
inc$dec <- pmin(10L, floor(inc$cw / sum(inc$WPFINWGT) * 10) + 1L)
inc$mw  <- ifelse(is.na(inc$match_per_worker_num), 0, inc$match_per_worker_num) * inc$WPFINWGT
dec_tbl <- inc |>
  group_by(dec) |>
  summarise(dollars = sum(mw), .groups = "drop") |>
  mutate(share = dollars / sum(dollars))
bottom3_share  <- sum(dec_tbl$share[dec_tbl$dec <= 3])
bottom_half_sh <- sum(dec_tbl$share[dec_tbl$dec <= 5])
zero_top_n     <- sum(dec_tbl$share < 0.001)

## ---- Automatic $100 seed contribution (scenario add-on) ----------------------
# seed_per_worker_num is $100 for every eligible worker and $0 otherwise (04_03),
# so the seed's incidence uses the same full-universe deciles as the match.
inc$sw <- inc$seed_per_worker_num * inc$WPFINWGT
seed_dec_tbl <- inc |>
  group_by(dec) |>
  summarise(seed_dollars = sum(sw), .groups = "drop") |>
  mutate(seed_share = seed_dollars / sum(seed_dollars))
seed_bottom3_share <- sum(seed_dec_tbl$seed_share[seed_dec_tbl$dec <= 3])
seed_cost_M_num    <- headline$seed_cost_M_num           # invariant across scenarios
seed_reach_M_num   <- headline$seed_recipients_M_num     # = eligible count
seed_only_M_num    <- headline$eligible_count_M_num - headline$participant_count_M_num

## Phased seed variants (04_03 Section 4b; participation-invariant)
sv_path <- file.path(dp, "seed_variants.parquet")
sv_avail <- file.exists(sv_path)
if (sv_avail) {
  sv <- as.data.frame(read_parquet(sv_path))
  sv_get <- function(nm) sv[sv$seed_variant_chr == nm, , drop = FALSE][1, ]
  sv_flat <- sv_get("flat"); sv_pro <- sv_get("pro_rata")
  sv_tap  <- sv_get("flat_then_taper"); sv_ext <- sv_get("extended_taper")
}

## ---- Routing split among eligible workers -----------------------------------
elig <- sim[sim$eligible_flag %in% TRUE, ]
route_tbl <- elig |>
  group_by(route_chr) |>
  summarise(
    elig_M    = sum(WPFINWGT, na.rm = TRUE) / 1e6,
    mean_magi = sum(magi_num * WPFINWGT, na.rm = TRUE) / sum(WPFINWGT, na.rm = TRUE),
    mean_rate = sum(match_rate_pp_num * WPFINWGT, na.rm = TRUE) / sum(WPFINWGT, na.rm = TRUE),
    cost_B    = sum(match_per_worker_num * WPFINWGT, na.rm = TRUE) / 1e9,
    .groups = "drop"
  )
route_lab <- c(employer_plan = "Existing workplace plan",
               universal_account = "New universal account")
route_tbl$label <- unname(route_lab[route_tbl$route_chr])
route_tbl <- route_tbl[order(-route_tbl$elig_M), ]

## ---- Generosity anchors for a single filer ----------------------------------
single_pivot    <- pv("single_mfs", "pivot_num")
single_endpoint <- pv("single_mfs", "endpoint_num")
single_median   <- pv("single_mfs", "data_median_magi_num")   # SIPP in-universe median (diagnostic)
# The policy anchor is the IRS all-single-filer median; the pivot is set at 0.8x
# it (the "75% rate at two-thirds the median" rule), so recover it for reference.
irs_single_median <- single_pivot / 0.8
# Schedule: rate(M) = 200 - (150 / pivot) * M (percent). Required own contribution
# to capture the full $1,000 credit = $1,000 / (rate / 100). The helper returns NA
# where the rate is non-positive (at/above the endpoint) so the cap-divide never
# yields Inf; the anchors used below all sit at strictly positive rates.
rate_at <- function(M) pmax(0, 200 - (150 / single_pivot) * M)
magi_at_rate <- function(r) (200 - r) / (150 / single_pivot)  # invert for r in pp
contrib_to_cap <- function(rate_pp) ifelse(rate_pp > 0, 1000 / (rate_pp / 100), NA_real_)
m_at_100 <- magi_at_rate(100)               # MAGI where the match rate is 100 percent
contrib_at_0   <- contrib_to_cap(rate_at(0))         # = $500 at the 200 percent floor
contrib_at_100 <- contrib_to_cap(rate_at(m_at_100))  # = $1,000
share_at_100   <- contrib_at_100 / m_at_100

## ---- Formatting helpers ------------------------------------------------------
d0  <- function(x) paste0("$", formatC(round(x), format = "d", big.mark = ","))
bil <- function(m) sprintf("$%.1f billion", m / 1000)          # input in $ millions
mil <- function(x) sprintf("%.1f million", x)                   # input in millions
pct <- function(frac, dig = 0) sprintf(paste0("%.", dig, "f percent"), 100 * frac)
ppt <- function(pp) sprintf("%.0f percent", pp)                 # a value already in pp

elig_share <- headline$eligible_count_M_num / universe_M
```

<style>
body{font-family:"Open Sans","Segoe UI",Helvetica,Arial,sans-serif;max-width:820px;
  margin:36px auto;padding:0 24px;color:#1a1a1a;line-height:1.5;font-size:15px;}
h1{font-size:26px;margin-bottom:2px;color:#024140;}
h2{font-size:18px;margin-top:30px;color:#19644d;border-bottom:2px solid #19644d;padding-bottom:3px;}
h3{font-size:15px;margin-top:20px;color:#19644d;}
.subtitle{font-size:15px;color:#555;}
.lede{font-size:16px;font-weight:600;color:#024140;margin:14px 0;}
.why{background:#FEECD6;border-left:4px solid #E1AD28;padding:10px 14px;margin:14px 0;border-radius:3px;}
table{border-collapse:collapse;width:100%;margin:14px 0;font-size:14px;}
th{background:#19644d;color:#fff;text-align:left;padding:6px 9px;}
td{padding:6px 9px;border-bottom:1px solid #e0e0e0;}
tr:nth-child(even) td{background:#f7f9f8;}
img{max-width:100%;height:auto;margin:8px 0;}
.cap{font-size:12px;color:#666;text-align:center;margin-top:-2px;margin-bottom:16px;}
.foot{font-size:12px;color:#777;margin-top:28px;border-top:1px solid #ddd;padding-top:10px;}
</style>

<p class="lede">A universal, automatic retirement account paired with a redesigned federal
match would extend a positive savings subsidy to roughly `r mil(headline$eligible_count_M_num)`
low- and moderate-income workers. Under automatic enrollment — the design's central assumption —
about `r mil(headline$participant_count_M_num)` would participate, at an estimated federal cost of
about `r bil(headline$annual_cost_M_num)` per year.</p>

<div class="why">
**Why it matters.** The savings match enacted in federal law rewards only workers who already
hold a workplace retirement account and who contribute enough to claim it. The binding flaw is not
the size of the match but its delivery: it has nowhere to land for a worker who has no account, so
the lowest earners — independent contractors, gig workers, and employees of firms that offer
nothing — gain the least. Roughly three-fifths of the dollars the enacted match would otherwise
direct to the bottom never arrive, because those workers lack the account the match must be
deposited into. The design described here closes that gap by pairing the match with a savings
account that every worker can use, raising the subsidy rate sharply at the bottom of the income
scale, and lowering the amount a worker must set aside to receive the full credit.
</div>

## What the policy does

The proposal has two moving parts, and the first does most of the work. Every covered worker is
automatically enrolled — at a default contribution of three percent of earnings — into a retirement
account. Workers already covered by a qualifying workplace plan keep that plan; everyone else,
including the self-employed and independent and platform workers, is routed into a new universal
account modeled on a low-cost, portable federal design. The self-employed, who have no payroll to
deduct from, reach the account through their estimated-tax payments or through the platforms they
work on, under a safe harbor so that offering the benefit does not reclassify a contractor as an
employee. This automatic account is what closes the access gap; it is the active ingredient, and
the bulk of the program's reach comes from supplying an account and a default contribution rather
than from the size of the subsidy.

Second, the federal match itself is redesigned. Rather than a flat 50 percent match that stops at a
fixed income cutoff, the match rate becomes a single declining line that depends on a worker's
income and tax-filing status. The rate begins at 200 percent for the lowest earners — one dollar of
federal credit for every fifty cents saved — and falls steadily as income rises, reaching zero at
the top of each eligibility band. The federal credit is capped at `r d0(1000)` per individual, so a
married couple in which both spouses work can receive up to `r d0(2000)` between them; because the
schedule is set by joint income but the cap applies per person, two equal earners are treated the
same married or single.

Two design choices keep the match simple to administer and easy on work incentives. The match is
computed on a worker's *prior-year* income, so the amount is known when the contribution is made,
requires no year-end reconciliation or clawback, and does not shrink because a worker earned a
raise during the year. And a worker's own contributions remain accessible — they can be withdrawn
without penalty, as under a Roth account — so enrolling does not lock away savings a low-income
household may need in an emergency; only the federal match carries a short holding period before it
vests.

```{r lens01-fig, out.width="92%"}
lens01_png <- file.path(fig_dir, "lens_01_match_rate.png")
if (file.exists(lens01_png)) knitr::include_graphics(lens01_png)
```
<p class="cap">The redesigned match rate (solid green) against the current-law Saver's Match (dashed
gold), for a single filer. The current match is a flat 50 percent that ends at a low fixed cutoff;
the redesign starts at 200 percent for the lowest earners and reaches far further up the income
scale.</p>

### Where the eligibility band falls

Each filing status has a *pivot* — the income level at which the match rate has declined to 50
percent. Below the pivot the rate is higher (rising to 200 percent at zero income); above it the rate
keeps falling along the same line until it reaches zero, where the match ends. The same geometry
applies to all three filing groups, scaled by the statutory ratios the enacted Saver's Match already
uses (head-of-household and joint bands are 1.5 and 2.0 times the single band).

```{r schedule-allfig, out.width="96%"}
sched_png <- file.path(fig_dir, "match_rate_by_magi.png")
if (file.exists(sched_png)) knitr::include_graphics(sched_png)
```
<p class="cap">The redesigned match rate (solid) against the current-law Saver's Match (dashed), by
filing status. The redesigned rate is far higher at low incomes and reaches more of the
distribution in every filing group.</p>

Rather than the fixed dollar cutoffs of current law — which Congress must legislate upward as wages
rise, and which otherwise let coverage erode — the band is anchored to administrative income data:
the single-filer pivot is set from the IRS median income for single filers (at two-thirds of that
median), indexed each year to chained-CPI inflation, and re-anchored to fresh IRS data about once a
decade. This keeps the policy on a transparent, replicable basis that a budget agency can reproduce,
and keeps the eligibility band tracking incomes without repeated acts of Congress.

```{r schedule-tbl}
sched <- data.frame(
  `Filing status` = c("Single / married filing separately", "Head of household", "Married filing jointly"),
  `Highest-match income` = c(d0(0), d0(0), d0(0)),
  `Income at a 50% rate (pivot)` = c(d0(pv("single_mfs","pivot_num")), d0(pv("hoh","pivot_num")), d0(pv("mfj","pivot_num"))),
  `Income where match ends`      = c(d0(pv("single_mfs","endpoint_num")), d0(pv("hoh","endpoint_num")), d0(pv("mfj","endpoint_num"))),
  check.names = FALSE
)
sched$`Highest-match income` <- "$0 (200% rate)"
kable(sched, align = c("l","l","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

## Who it reaches

The estimates rest on a large, nationally representative U.S. household survey that records each
person's earnings, income, household composition, and access to a workplace retirement plan.
Survey weights are applied throughout so that every figure reflects the national population rather
than the sample count.

The analysis begins with the roughly `r mil(adults_18p_M)` U.S. residents age 18 or older and
narrows in steps to the working population the policy is built for: of those adults, about
`r mil(earners_M)` have earnings, and after setting aside dependents and full-time students and
keeping only those with a determinable tax-filing status, a universe of about `r mil(universe_M)`
workers remains. Within it, an estimated `r mil(headline$eligible_count_M_num)` workers — close to
`r pct(elig_share)` of the universe — fall inside the redesigned eligibility band and would qualify
for a positive match.

Eligible workers reach the match through one of two channels. Those who already hold a workplace
plan claim it there; the rest reach it through the new universal account. The universal account
carries the larger share of eligible workers, both because it absorbs self-employed workers and
because workers without a plan at work tend to have lower incomes — and therefore higher match
rates.

```{r route-tbl}
rt <- data.frame(
  `Channel` = route_tbl$label,
  `Eligible workers` = mil(route_tbl$elig_M),
  `Average income` = d0(route_tbl$mean_magi),
  `Average match rate` = paste0(round(route_tbl$mean_rate), "%"),
  check.names = FALSE
)
kable(rt, align = c("l","r","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

The federal dollars land where the need is greatest. Of the full-participation match, an estimated
`r pct(bottom3_share)` flows to the lowest-earning 30 percent of the workforce (the bottom three
income deciles), and nearly all of it to the lower-earning half. The highest-income workers receive
nothing, because the match phases out entirely before their income range begins.

```{r incidence-fig, out.width="74%"}
inc_png <- file.path(fig_dir, "incidence_by_decile.png")
if (file.exists(inc_png)) knitr::include_graphics(inc_png)
```
<p class="cap">Share of federal match dollars by income decile across the full workforce (1 = lowest).
The subsidy is weighted toward lower-income savers; higher deciles receive little or nothing.</p>

## How generous it is

The redesign is most visible at the bottom of the income scale, where the current match does the
least. Three measures make the point, using a single filer as the example.

- **A far higher subsidy rate.** At very low earnings the match rate is 200 percent, meaning a
  worker who sets aside just `r d0(contrib_at_0)` receives the full `r d0(1000)` federal credit. The
  rate is still 100 percent at about `r d0(m_at_100)` of income, where a `r d0(contrib_at_100)`
  contribution — roughly `r pct(share_at_100, 1)` of earnings — captures the full credit.
- **An easier path to the maximum credit.** Under current law a worker must contribute the full
  `r d0(2000)` to receive the `r d0(1000)` maximum. Under the redesign, low earners reach the same
  maximum on a fraction of that amount, which matters because lower-income workers were never
  likely to set aside `r d0(2000)`.
- **A larger average award.** Across all participants the redesign delivers an average federal
  match of about `r d0(headline$avg_match_per_person_num)` per year`r if (cl_avail) paste0(", set against roughly ", d0(cl_avg), " per worker under the current-law match at full participation in the same population") else ""`.

The high rate at the bottom is a deliberate choice. A dollar is hardest to move out of day-to-day
consumption for the lowest earners, so the schedule over-matches there — a 200 percent rate — to make
a small contribution worth enough to pull a household into saving at all. The two figures below
trace what that means for the worker's side of the bargain: how much a single filer must set aside
to capture the full `r d0(1000)` credit, in dollars and as a share of earnings.

```{r lens02-fig, out.width="92%"}
lens02_png <- file.path(fig_dir, "lens_02_required_contribution.png")
if (file.exists(lens02_png)) knitr::include_graphics(lens02_png)
```
<p class="cap">The out-of-pocket contribution needed to receive the full `r d0(1000)` match. At the
bottom of the band just `r d0(contrib_at_0)` unlocks it; the required amount rises with income to
`r d0(2000)` at the pivot and then climbs out of reach as the match rate falls toward zero. The
current-law match (gold) requires the full `r d0(2000)` and only below a low fixed cutoff.</p>

```{r lens03-fig, out.width="92%"}
lens03_png <- file.path(fig_dir, "lens_03_required_share_of_earnings.png")
if (file.exists(lens03_png)) knitr::include_graphics(lens03_png)
```
<p class="cap">The same requirement as a share of earnings is U-shaped, bottoming at about
`r pct(share_at_100, 1)` of earnings near `r d0(m_at_100)` of income — just above the three percent
automatic default, so a worker who nudges their contribution a little above the default captures the
full credit across the middle of the band. Under current law (gold) the cap is reachable only at the
very bottom.</p>

## What it costs

The central estimate is an annual federal cost of about `r bil(headline$annual_cost_M_num)`,
reaching `r mil(headline$participant_count_M_num)` participating savers at an automatic-enrollment
participation rate of `r pct(headline$takeup_rate_num)`. Participation is the single largest driver
of cost, so the estimate is best read as a range. The lower bound applies the participation rate
observed today among workers who already have plan access (about `r pct(obs_cond$takeup_rate_num)`) —
a figure the evidence on automatic enrollment suggests understates how many workers stay in a sticky
payroll default, which is why the `r pct(headline$takeup_rate_num)` auto-enrollment case is taken as
the headline rather than the floor. The upper bound assumes every eligible worker participates.

```{r cost-tbl}
cost <- data.frame(
  `Participation scenario` = c(
    "Observed participation behavior (lower bound)",
    "Automatic enrollment at 80 percent (headline)",
    "Full participation (upper bound)"),
  `Take-up` = c(pct(obs_cond$takeup_rate_num),
                pct(headline$takeup_rate_num), pct(s_full$takeup_rate_num)),
  `Participants` = c(mil(obs_cond$participant_count_M_num),
                     mil(headline$participant_count_M_num), mil(s_full$participant_count_M_num)),
  `Annual cost` = c(bil(obs_cond$annual_cost_M_num),
                    bil(headline$annual_cost_M_num), bil(s_full$annual_cost_M_num)),
  check.names = FALSE
)
kable(cost, align = c("l","r","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

Two features of the cost range stand out. The structural ceiling — the cost if every eligible worker
participated — is about `r bil(s_full$annual_cost_M_num)`, so the program cannot cost more than that
at any plausible behavioral response. And because the match is paid only on the contributions
workers actually make, cost tracks participation closely: the program's reach, and its price, both
depend on the automatic account doing its job. These figures are gross transfer costs. Over a longer
horizon part of that cost is recovered: workers who reach retirement with more savings draw less on
means-tested programs such as Medicaid and Supplemental Security Income, an offset a budget-model
analysis of the closely related Retirement Savings for Americans Act estimated grows large enough to
turn that proposal net-positive after several decades. The cap on the per-person credit here bounds
the gross cost well below that proposal's, so the same direction of offset should apply at lower
expense.

```{r contrast, eval=cl_avail}
cat("### How it compares with current law\n\n")
contrast <- data.frame(
  ` ` = c("Match rate at the bottom", "Income thresholds",
          "Eligible workers", "Federal cost", "Average match per worker"),
  `Current-law Saver's Match` = c(
    "50%",
    "Fixed dollar cutoffs",
    mil(cl_elig_M),
    bil(cl_cost_M),
    d0(cl_avg)),
  `This proposal` = c(
    "Up to 200%",
    "Anchored to the IRS median by filing status",
    mil(headline$eligible_count_M_num),
    bil(s_full$annual_cost_M_num),
    d0(s_full$avg_match_per_person_num)),
  check.names = FALSE
)
print(kable(contrast, align = c("l","l","l"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed")))
cat("\n\nBoth columns apply the same working population at full participation, so the difference reflects design alone — a higher subsidy rate at the bottom, an income-anchored band reaching more workers, and a larger average award.\n")
```

## Scenario add-on: an automatic $100 seed contribution

A match, however generous, reaches only workers who contribute. As a scenario extension, the model
adds an automatic `r d0(100)` annual federal contribution deposited into the account of every
eligible worker — whether or not the worker contributes anything of their own. The seed does not
count toward the `r d0(1000)` match cap, and because it is unconditional its cost does not depend
on the participation assumption.

Three results summarize the scenario:

- **Reach.** The seed reaches all `r mil(seed_reach_M_num)` eligible workers in every scenario. Under
  the headline 80 percent auto-enrollment case, that includes about `r mil(seed_only_M_num)` eligible
  workers who would not be contributing and would therefore receive no match at all — the seed is the
  only federal retirement dollar that reaches them, and it arrives in an account that then exists to
  receive a tax refund, a windfall, or a future raise.
- **Cost.** The seed adds about `r bil(seed_cost_M_num)` per year at any participation rate, raising
  the headline total from `r bil(headline$annual_cost_M_num)` to
  `r bil(headline$total_cost_incl_seed_M_num)` and the full-participation ceiling to
  `r bil(s_full$total_cost_incl_seed_M_num)`.
- **Distribution.** Because the seed is a flat amount per person, its dollars follow the eligible
  population rather than contribution size: an estimated `r pct(seed_bottom3_share)` of seed dollars
  flow to the lowest-earning 30 percent of the workforce, and the seed is relatively most important
  exactly where match dollars are smallest — for workers whose contributions, and therefore matches,
  are near zero.

```{r seed-cost-tbl}
seed_cost_tbl <- data.frame(
  `Participation scenario` = c(
    "Observed participation behavior (lower bound)",
    "Automatic enrollment at 80 percent (headline)",
    "Full participation (upper bound)"),
  `Match cost` = c(bil(obs_cond$annual_cost_M_num),
                   bil(headline$annual_cost_M_num), bil(s_full$annual_cost_M_num)),
  `Seed cost` = c(bil(obs_cond$seed_cost_M_num),
                  bil(headline$seed_cost_M_num), bil(s_full$seed_cost_M_num)),
  `Total including seed` = c(bil(obs_cond$total_cost_incl_seed_M_num),
                             bil(headline$total_cost_incl_seed_M_num),
                             bil(s_full$total_cost_incl_seed_M_num)),
  check.names = FALSE
)
kable(seed_cost_tbl, align = c("l","r","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

```{r seed-incidence-fig, out.width="74%"}
seed_png <- file.path(fig_dir, "incidence_seed_vs_match.png")
if (file.exists(seed_png)) knitr::include_graphics(seed_png)
```
<p class="cap">Federal dollars by income decile across the full workforce, decomposing the
full-participation match (green) and the automatic `r d0(100)` seed (gold). The seed is small
against the match in dollar terms, but it is the only component that reaches eligible workers who
do not contribute.</p>

The case for the seed is not the size of the first-year benefit — `r d0(100)` is a modest sum — but
account formation: it guarantees that every eligible worker ends the year with a funded retirement
account, shifting the remaining policy problem from opening accounts to growing balances, which a
future Congress could address by raising the contribution. The caveat travels in the other
direction: an automatic deposit requires the universal account infrastructure to reach workers the
payroll system does not, so the seed's administrative burden is concentrated on exactly the workers
who are hardest to enroll.

### Phasing the seed instead of ending it abruptly

`r if (sv_avail) "" else "<!-- seed_variants.parquet not found; variant table suppressed -->"`

The flat design has one rough edge: a worker just inside the eligibility band receives the full
`r d0(100)` while a worker one dollar above it receives nothing. Three phased variants remove that
cliff in different ways, trading cost against how many workers keep the full amount. A *pro-rata*
seed scales with the match rate itself, declining from `r d0(100)` at zero income along the same
line as the match. A *flat-then-taper* seed pays the full `r d0(100)` up to the pivot — the income
where the match rate has fallen to 50 percent — and then declines to zero at the match endpoint,
mirroring the geometry of the current-law Saver's Match. An *extended-taper* seed pays the full
`r d0(100)` across the entire eligible band and phases out beyond it, so no eligible worker loses
anything relative to the flat design and workers just above the band receive a partial seed.

```{r seed-variant-tbl, eval=sv_avail}
sv_row <- function(s, label) data.frame(
  `Seed design` = label,
  `Annual cost` = bil(s$seed_cost_M_num),
  `Recipients` = mil(s$seed_recipients_M_num),
  `Receive the full $100` = if (s$seed_full_amount_M_num < 0.5) "—" else mil(s$seed_full_amount_M_num),
  `Share of seed dollars to the bottom 30%` = pct(s$seed_bottom3_share_num),
  check.names = FALSE
)
sv_tbl <- rbind(
  sv_row(sv_flat, "Flat $100 (headline)"),
  sv_row(sv_pro,  "Pro-rata with match rate"),
  sv_row(sv_tap,  "Flat, then taper"),
  sv_row(sv_ext,  "Extended taper")
)
kable(sv_tbl, align = c("l","r","r","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

```{r seed-variant-fig, out.width="74%"}
sv_png <- file.path(fig_dir, "seed_variant_schedules.png")
if (file.exists(sv_png)) knitr::include_graphics(sv_png)
```
<p class="cap">The four seed designs for a single filer. The flat design stops abruptly at the match
endpoint; the phased variants remove the cliff by declining inside the band (pro-rata,
flat-then-taper) or by extending past it (extended taper).</p>

The choice among them is a policy dial rather than a modeling question. The pro-rata design is the
cheapest`r if (sv_avail) paste0(" (", bil(sv_pro$seed_cost_M_num), ")") else ""` and the most
tightly targeted, but it undercuts the account-seeding rationale — most recipients get a deposit
well short of `r d0(100)`. The flat-then-taper design preserves the full seed for the lower half of
the band at modest savings. The extended taper is the most generous`r if (sv_avail) paste0(" (",
bil(sv_ext$seed_cost_M_num), ", reaching ", mil(sv_ext$seed_recipients_M_num), " workers)") else ""`
and treats the boundary problem by reaching past it rather than shrinking inside it.

## How the estimates were produced, and what to keep in mind

The cost and reach figures come from a microsimulation: the policy rules are applied to each worker
in the survey, the federal match is computed for that worker given their income, filing status, and
savings channel, and the results are summed using national population weights. Incomes are projected
forward to the policy's first effective tax year using a standard wage-growth assumption so that the
dollar thresholds line up with the year the policy would take effect.

A few assumptions shape the numbers and should travel with them. The estimates are descriptive: they
show what the policy rules would cost and whom they would reach under stated participation
assumptions, not a behavioral forecast of how saving or work decisions might change. Participation is
the largest source of uncertainty, which is why the cost is reported as a range; and the federal
match is paid only on contributions workers actually make, so the estimates assume the automatic
three percent default is contributed. They also do not model how employers might respond — in
particular, whether some small firms would drop their own plans once a subsidized universal account
exists, a displacement risk that independent analyses of similar federal-account proposals have
flagged and that would reduce net new saving. Because the underlying data are a sample of
households, the figures are point estimates rather than exact counts and carry no statistical
confidence intervals, so small differences should not be over-interpreted. The eligibility band is
anchored to administrative IRS income data, but each worker's income is measured in a household
survey that approximates — and is drawn from a somewhat different population than — the tax records a
revenue agency would use, so a tax-administered version would shift the counts modestly; the survey's
incomes are aged forward to the policy's first year for dollar comparability, but the population's
composition and behavior are not re-projected. Finally, the design's withdrawal rules follow the
portable-account model of pending legislation — a worker's own contributions stay accessible, while
the federal match must season briefly before it vests — but the leakage those rules permit, the cost
of administering the universal account, and the standing up of that account are not separately
modeled here; each matters for an implementing statute more than for the headline transfer cost.

<p class="foot">
Source: Author's analysis of a nationally representative U.S. household survey (2024), projected to
the policy's first effective tax year. All population figures are survey-weighted. Costs are annual
and expressed in current dollars. Figures reflect the most recent model run and may update as inputs
are refreshed.
</p>
