---
title: "A Universal Saver's Match"
subtitle: "Design, reach, generosity, and cost of a redesigned federal retirement-savings match"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
params:
  root: NULL
output:
  html_document:
    self_contained: true
    theme: null
    highlight: null
    mathjax: null
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      results = "asis", fig.align = "center")

suppressPackageStartupMessages({
  library(arrow)
  library(dplyr)
  library(readr)
  library(knitr)
  library(kableExtra)
})

## ---- Resolve the project root (param -> env var -> upward search) -----------
root <- params$root
if (is.null(root) || !nzchar(root) || !dir.exists(root)) {
  root <- Sys.getenv("EIG_PROJECT_ROOT", unset = "")
}
if (!nzchar(root) || !dir.exists(root)) {
  cand <- normalizePath(getwd())
  while (!dir.exists(file.path(cand, "Infrastructure")) && cand != dirname(cand)) {
    cand <- dirname(cand)
  }
  root <- cand
}

dp <- file.path(root, "data", "processed", "universal_sm_hybrid")
rep_dir <- file.path(root, "output", "reports", "universal_sm_hybrid")
fig_dir <- file.path(root, "output", "figures", "universal_sm_hybrid")

## ---- Load the live pipeline artifacts ---------------------------------------
scen   <- as.data.frame(read_parquet(file.path(dp, "scenario_results.parquet")))
pivots <- readRDS(file.path(dp, "pivot_table.rds"))$pivot_tbl
sim    <- as.data.frame(read_parquet(file.path(dp, "simulation_results.parquet")))
funnel <- read_csv(file.path(rep_dir, "universe_funnel.csv"), show_col_types = FALSE)

get_scen <- function(nm) scen[scen$scenario_name_chr == nm, , drop = FALSE][1, ]
headline <- get_scen("headline_sipp_observed_conditional")
dc_route <- get_scen("headline_dc_access_row_level")
uv_route <- get_scen("headline_universal_account_uniform")
s_noauto <- get_scen("sens_no_auto_5_7pct")
s_auto80 <- get_scen("sens_auto_enroll_80pct")
s_full   <- get_scen("sens_full_participation_100pct")

pv <- function(grp, col) pivots[[col]][pivots$filing_group_chr == grp]

## ---- Current-law Saver's Match contrast (same population, full take-up) ------
cl_avail <- FALSE
cl_path <- file.path(root, "output", "tables", "main", "sm_jct_replication_scenarios.xlsx")
if (requireNamespace("openxlsx", quietly = TRUE) && file.exists(cl_path)) {
  cl <- openxlsx::read.xlsx(cl_path, sheet = "All Scenario Results")
  cl_row <- cl[cl$Scenario == "universal_m100", , drop = FALSE]
  if (nrow(cl_row) == 1L) {
    cl_avail   <- TRUE
    cl_elig_M  <- cl_row[["Eligible.(M)"]]
    cl_cost_M  <- cl_row[["Annual.Cost.($M)"]]
    cl_avg     <- cl_row[["Avg.Match/Person.($)"]]
  }
}

## ---- Universe / sample funnel -----------------------------------------------
fn <- function(step) funnel$weighted_n_num[funnel$step_label_chr == step] / 1e6
adults_18p_M <- fn("02_age_18_plus")
earners_M    <- fn("05_has_annual_earned_income")
universe_M   <- fn("06_valid_filing_group")

## ---- Distributional incidence (full-participation match dollars) ------------
inc <- sim[!is.na(sim$magi_num), ]
inc <- inc[order(inc$magi_num), ]
inc$cw  <- cumsum(inc$WPFINWGT)
inc$dec <- pmin(10L, floor(inc$cw / sum(inc$WPFINWGT) * 10) + 1L)
inc$mw  <- ifelse(is.na(inc$match_per_worker_num), 0, inc$match_per_worker_num) * inc$WPFINWGT
dec_tbl <- inc |>
  group_by(dec) |>
  summarise(dollars = sum(mw), .groups = "drop") |>
  mutate(share = dollars / sum(dollars))
bottom3_share  <- sum(dec_tbl$share[dec_tbl$dec <= 3])
bottom_half_sh <- sum(dec_tbl$share[dec_tbl$dec <= 5])
zero_top_n     <- sum(dec_tbl$share < 0.001)

## ---- Routing split among eligible workers -----------------------------------
elig <- sim[sim$eligible_flag %in% TRUE, ]
route_tbl <- elig |>
  group_by(route_chr) |>
  summarise(
    elig_M    = sum(WPFINWGT, na.rm = TRUE) / 1e6,
    mean_magi = sum(magi_num * WPFINWGT, na.rm = TRUE) / sum(WPFINWGT, na.rm = TRUE),
    mean_rate = sum(match_rate_pp_num * WPFINWGT, na.rm = TRUE) / sum(WPFINWGT, na.rm = TRUE),
    cost_B    = sum(match_per_worker_num * WPFINWGT, na.rm = TRUE) / 1e9,
    .groups = "drop"
  )
route_lab <- c(employer_plan = "Existing workplace plan",
               universal_account = "New universal account")
route_tbl$label <- unname(route_lab[route_tbl$route_chr])
route_tbl <- route_tbl[order(-route_tbl$elig_M), ]

## ---- Generosity anchors for a single filer ----------------------------------
single_pivot    <- pv("single_mfs", "pivot_num")
single_endpoint <- pv("single_mfs", "endpoint_num")
single_median   <- pv("single_mfs", "data_median_magi_num")
# Schedule: rate(M) = 200 - (150 / pivot) * M (percent). Required own contribution
# to capture the full $1,000 credit = $1,000 / (rate / 100). The helper returns NA
# where the rate is non-positive (at/above the endpoint) so the cap-divide never
# yields Inf; the anchors used below all sit at strictly positive rates.
rate_at <- function(M) pmax(0, 200 - (150 / single_pivot) * M)
magi_at_rate <- function(r) (200 - r) / (150 / single_pivot)  # invert for r in pp
contrib_to_cap <- function(rate_pp) ifelse(rate_pp > 0, 1000 / (rate_pp / 100), NA_real_)
m_at_100 <- magi_at_rate(100)               # MAGI where the match rate is 100 percent
contrib_at_0   <- contrib_to_cap(rate_at(0))         # = $500 at the 200 percent floor
contrib_at_100 <- contrib_to_cap(rate_at(m_at_100))  # = $1,000
share_at_100   <- contrib_at_100 / m_at_100

## ---- Formatting helpers ------------------------------------------------------
d0  <- function(x) paste0("$", formatC(round(x), format = "d", big.mark = ","))
bil <- function(m) sprintf("$%.1f billion", m / 1000)          # input in $ millions
mil <- function(x) sprintf("%.1f million", x)                   # input in millions
pct <- function(frac, dig = 0) sprintf(paste0("%.", dig, "f percent"), 100 * frac)
ppt <- function(pp) sprintf("%.0f percent", pp)                 # a value already in pp

elig_share <- headline$eligible_count_M_num / universe_M
```

<style>
body{font-family:"Open Sans","Segoe UI",Helvetica,Arial,sans-serif;max-width:820px;
  margin:36px auto;padding:0 24px;color:#1a1a1a;line-height:1.5;font-size:15px;}
h1{font-size:26px;margin-bottom:2px;color:#024140;}
h2{font-size:18px;margin-top:30px;color:#19644d;border-bottom:2px solid #19644d;padding-bottom:3px;}
h3{font-size:15px;margin-top:20px;color:#19644d;}
.subtitle{font-size:15px;color:#555;}
.lede{font-size:16px;font-weight:600;color:#024140;margin:14px 0;}
.why{background:#FEECD6;border-left:4px solid #E1AD28;padding:10px 14px;margin:14px 0;border-radius:3px;}
table{border-collapse:collapse;width:100%;margin:14px 0;font-size:14px;}
th{background:#19644d;color:#fff;text-align:left;padding:6px 9px;}
td{padding:6px 9px;border-bottom:1px solid #e0e0e0;}
tr:nth-child(even) td{background:#f7f9f8;}
img{max-width:100%;height:auto;margin:8px 0;}
.cap{font-size:12px;color:#666;text-align:center;margin-top:-2px;margin-bottom:16px;}
.foot{font-size:12px;color:#777;margin-top:28px;border-top:1px solid #ddd;padding-top:10px;}
</style>

<p class="lede">A universal, automatic retirement account paired with a redesigned federal
match would extend a meaningful savings subsidy to roughly `r mil(headline$eligible_count_M_num)`
low- and moderate-income workers at an estimated headline cost of about `r bil(headline$annual_cost_M_num)` per year.</p>

<div class="why">
**Why it matters.** The savings match enacted in federal law rewards only workers who already
hold a workplace retirement account and who contribute enough to claim it. Workers without a plan
at work — independent contractors, gig workers, and employees of firms that offer nothing — are
left out, and the lowest earners gain the least. The design described here pairs the match with a
savings account that every worker can use, raises the subsidy rate sharply at the bottom of the
income scale, and lowers the amount a worker must set aside to receive the full credit.
</div>

## What the policy does

The proposal has two moving parts. First, every covered worker is automatically enrolled — at a
default contribution of three percent of earnings — into a retirement account. Workers already
covered by a qualifying workplace plan keep that plan; everyone else is routed into a new universal
account modeled on a low-cost, portable federal design. This closes the access gap that keeps the
current match from reaching the workers it was meant to help.

Second, the federal match itself is redesigned. Rather than a flat 50 percent match that stops at a
fixed income cutoff, the match rate becomes a single declining line that depends on a worker's
income and tax-filing status. The rate begins at 200 percent for the lowest earners — three federal
dollars for every one the worker saves becomes one dollar of credit for every fifty cents
saved — and falls steadily as income rises, reaching zero at the top of each eligibility band. The
federal credit is capped at `r d0(1000)` per individual, so a married couple in which both spouses
work can receive up to `r d0(2000)` between them.

Because the schedule is anchored to the income distribution rather than to fixed dollar thresholds,
the eligibility band scales automatically with filing status and can be re-anchored as incomes
change over time.

```{r schedule-fig, out.width="86%"}
sched_png <- file.path(fig_dir, "match_rate_by_magi.png")
if (file.exists(sched_png)) knitr::include_graphics(sched_png)
```
<p class="cap">The redesigned match rate (solid) against the current-law Saver's Match (dashed),
by filing status. The redesigned rate is far higher at low incomes and reaches more of the
distribution.</p>

### Where the eligibility band falls

```{r schedule-tbl}
sched <- data.frame(
  `Filing status` = c("Single / married filing separately", "Head of household", "Married filing jointly"),
  `Highest-match income` = c(d0(0), d0(0), d0(0)),
  `Income at a 50% rate (pivot)` = c(d0(pv("single_mfs","pivot_num")), d0(pv("hoh","pivot_num")), d0(pv("mfj","pivot_num"))),
  `Income where match ends`      = c(d0(pv("single_mfs","endpoint_num")), d0(pv("hoh","endpoint_num")), d0(pv("mfj","endpoint_num"))),
  check.names = FALSE
)
sched$`Highest-match income` <- "$0 (200% rate)"
kable(sched, align = c("l","l","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

## Who it reaches

The estimates rest on a large, nationally representative U.S. household survey that records each
person's earnings, income, household composition, and access to a workplace retirement plan.
Survey weights are applied throughout so that every figure reflects the national population rather
than the sample count.

The analysis begins with the roughly `r mil(adults_18p_M)` U.S. residents age 18 or older and
narrows to the working population most relevant to the policy: people with earnings and a
determinable tax-filing status, after setting aside dependents and full-time students. That leaves
a universe of about `r mil(universe_M)` workers. Within it, an estimated
`r mil(headline$eligible_count_M_num)` workers — close to `r pct(elig_share)` of the universe — fall
inside the redesigned eligibility band and would qualify for a positive match.

```{r funnel-fig, out.width="74%"}
fn_png <- file.path(fig_dir, "universe_funnel.png")
if (file.exists(fn_png)) knitr::include_graphics(fn_png)
```
<p class="cap">From all adults to the working population that the match can reach.</p>

Eligible workers reach the match through one of two channels. Those who already hold a workplace
plan claim it there; the rest reach it through the new universal account. The universal account
carries the larger share of eligible workers, both because it absorbs self-employed workers and
because workers without a plan at work tend to have lower incomes — and therefore higher match
rates.

```{r route-tbl}
rt <- data.frame(
  `Channel` = route_tbl$label,
  `Eligible workers` = mil(route_tbl$elig_M),
  `Average income` = d0(route_tbl$mean_magi),
  `Average match rate` = paste0(round(route_tbl$mean_rate), "%"),
  check.names = FALSE
)
kable(rt, align = c("l","r","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

## How generous it is

The redesign is most visible at the bottom of the income scale, where the current match does the
least. Three measures make the point, using a single filer as the example.

- **A far higher subsidy rate.** At very low earnings the match rate is 200 percent, meaning a
  worker who sets aside just `r d0(contrib_at_0)` receives the full `r d0(1000)` federal credit. The
  rate is still 100 percent at about `r d0(m_at_100)` of income, where a `r d0(contrib_at_100)`
  contribution — roughly `r pct(share_at_100, 1)` of earnings — captures the full credit.
- **An easier path to the maximum credit.** Under current law a worker must contribute the full
  `r d0(2000)` to receive the `r d0(1000)` maximum. Under the redesign, low earners reach the same
  maximum on a fraction of that amount, which matters because lower-income workers were never
  likely to set aside `r d0(2000)`.
- **A larger average award.** Across all participants the redesign delivers an average federal
  match of about `r d0(headline$avg_match_per_person_num)` per year`r if (cl_avail) paste0(", set against roughly ", d0(cl_avg), " per worker under the current-law match at full participation in the same population. The two averages rest on different participation assumptions, so read this as a broad comparison rather than a like-for-like one") else ""`.

The federal dollars are concentrated where the need is greatest. Of the full-participation match,
an estimated `r pct(bottom3_share)` flows to the lowest-earning 30 percent of the workforce (the
bottom three income deciles), and nearly all of it to the lower-earning half. The highest-income
workers receive nothing, because the match phases out entirely before their income range begins.

```{r incidence-fig, out.width="74%"}
inc_png <- file.path(fig_dir, "incidence_by_decile.png")
if (file.exists(inc_png)) knitr::include_graphics(inc_png)
```
<p class="cap">Share of federal match dollars by income decile across the full workforce (1 = lowest).
The subsidy is weighted toward lower-income savers; higher deciles receive little or nothing.</p>

## What it costs

The headline estimate is an annual federal cost of about `r bil(headline$annual_cost_M_num)`,
reaching `r mil(headline$participant_count_M_num)` participating savers at a take-up rate of about
`r pct(headline$takeup_rate_num)`. Cost depends heavily on how many eligible workers actually
participate, which is why the estimate is presented as a range rather than a single number.

```{r cost-tbl}
cost <- data.frame(
  `Participation scenario` = c(
    "Low take-up, no automatic enrollment",
    "Headline (observed participation behavior)",
    "Automatic enrollment at 80 percent participation",
    "Full participation (upper bound)"),
  `Take-up` = c(pct(s_noauto$takeup_rate_num), pct(headline$takeup_rate_num),
                pct(s_auto80$takeup_rate_num), pct(s_full$takeup_rate_num)),
  `Participants` = c(mil(s_noauto$participant_count_M_num), mil(headline$participant_count_M_num),
                     mil(s_auto80$participant_count_M_num), mil(s_full$participant_count_M_num)),
  `Annual cost` = c(bil(s_noauto$annual_cost_M_num), bil(headline$annual_cost_M_num),
                    bil(s_auto80$annual_cost_M_num), bil(s_full$annual_cost_M_num)),
  check.names = FALSE
)
kable(cost, align = c("l","r","r","r"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed"))
```

Two features of the cost range stand out. The structural ceiling — the cost if every eligible
worker participated fully — is about `r bil(s_full$annual_cost_M_num)`, so the program cannot cost
more than that at any plausible behavioral response. And the floor scenario, which assumes none of
the automatic-enrollment machinery works and participation matches the low claim rate of the
predecessor tax credit, would cost only about `r bil(s_noauto$annual_cost_M_num)` — a reminder that
the policy's value depends on the automatic account doing its job.

```{r contrast, eval=cl_avail}
cat("### How it compares with current law\n\n")
contrast <- data.frame(
  ` ` = c("Match rate at the bottom", "Income thresholds",
          "Eligible workers", "Federal cost", "Average match per worker"),
  `Current-law Saver's Match` = c(
    "50%",
    "Fixed dollar cutoffs",
    mil(cl_elig_M),
    bil(cl_cost_M),
    d0(cl_avg)),
  `This proposal` = c(
    "Up to 200%",
    "Anchored to the income distribution",
    mil(headline$eligible_count_M_num),
    paste0(bil(headline$annual_cost_M_num), " (headline)"),
    d0(headline$avg_match_per_person_num)),
  check.names = FALSE
)
print(kable(contrast, align = c("l","l","l"), row.names = FALSE) |>
  kable_styling(full_width = TRUE, bootstrap_options = c("condensed")))
cat("\n\nThe current-law comparison applies the same population and a full-participation assumption, so the difference reflects design alone.\n")
```

## How the estimates were produced, and what to keep in mind

The cost and reach figures come from a microsimulation: the policy rules are applied to each worker
in the survey, the federal match is computed for that worker given their income, filing status, and
savings channel, and the results are summed using national population weights. Incomes are projected
forward to the policy's first effective tax year using a standard wage-growth assumption so that the
dollar thresholds line up with the year the policy would take effect.

A few assumptions shape the numbers and should travel with them. The estimates are descriptive: they
show what the policy rules would cost and whom they would reach under stated participation
assumptions, not a behavioral forecast of how saving or work decisions might change. Participation is
the largest source of uncertainty, which is why the cost is reported as a range; and the federal
match is paid only on contributions workers actually make, so the estimates assume the automatic
three percent default is contributed. Because the underlying data are a sample of households, the
figures are point estimates rather than exact counts; they are not reported with statistical
confidence intervals, so small differences should not be over-interpreted. The income measure is
drawn from a household survey and approximates — but does not exactly equal — the income concept a
tax agency would use, so the
eligibility thresholds would shift modestly under a tax-administered version. The survey's incomes
are aged forward to the policy's first year for dollar comparability, but the population's
composition and saving behavior are not re-projected. The estimates also leave aside administrative
costs, the cost of running the universal account, and any rules governing early withdrawals; each
matters for an implementing statute but not for the headline transfer cost reported here.

<p class="foot">
Source: Author's analysis of a nationally representative U.S. household survey (2024), projected to
the policy's first effective tax year. All population figures are survey-weighted. Costs are annual
and expressed in current dollars. Figures reflect the most recent model run and may update as inputs
are refreshed.
</p>
